---
title: "sf Filters"
author: "Zach"
date: 2020-01-29
output:
  github_document
---

The [sf](https://r-spatial.github.io/sf/index.html) package has a wide variety
of tools, but the documentation at the function level is poor (the
[vignettes](https://r-spatial.github.io/sf/articles/sf3.html#geometrical-operations)
are nice, but cumbersome for reference).

The following are recipies for using sf tools to `filter()` GIS data.

```{r setup, echo=FALSE}
library(tidyverse)
library(sf)
```

### Running Example Data

The following simple data are from an [sf vignette](https://r-spatial.github.io/sf/articles/sf3.html#geometrical-operations):

```{r simple-data}
## Red set
b0 = st_polygon(list(rbind(c(-1,-1), c(1,-1), c(1,1), c(-1,1), c(-1,-1))))
b1 = b0 + 2
b2 = b0 + c(-0.2, 2)
x = st_sfc(b0, b1, b2)
sf_x <- st_sf(x)

## Green set
a0 = b0 * 0.8
a1 = a0 * 0.5 + c(2, 0.7)
a2 = a0 + 1
a3 = b0 * 0.5 + c(1.5, -0.5)
y = st_sfc(a0,a1,a2,a3)
sf_y <- st_sf(y)

ggplot() +
  geom_sf(data = sf_x, color = "red", fill = NA) +
  geom_sf(data = sf_y, color = "green", fill = NA) +
  theme_void()
```

## Binary Logical Operations
<!-- -------------------------------------------------- -->

The sf binary logical operations are based on
[DE-9IM](https://en.wikipedia.org/wiki/DE-9IM); a framework for defining binary
relationships between topological objects. The following `sf` calls implement
these binary relationships.

|  |  |
|--|--|
| st_intersects(x, y) | x intersects y |
| st_disjoint(x, y) | x not Intersects y |
| st_touches(x, y) | x meets y |
| st_crosses(x, y) | x crosses y |
| st_within(x, y) | x inside y |
| st_contains(x, y) | x contains y |
| st_overlaps(x, y) | x overlaps y |
| st_equals(x, y) | x equals y |
| st_covers(x, y) | x covers y |
| st_covered_by(x, y) | x covered by y |
| st_equals_exact(x, y) | x equals y exactly |

These functions---by default---return a sparse matrix of results:

```{r sparse-return}
y %>% st_intersects(., b0)
```

To use these functions with `filter()`, we'll need to disable the sparse return:

```{r dense-return}
y %>% st_intersects(., b0, sparse = FALSE)
```

We can easily integrate this in a tidyverse pipeline:

```{r pipeline-example}
sf_y %>% 
  filter(st_intersects(., b0, sparse = FALSE))
```


### Illustration

Here we illustrate all of the binary relationships listed above.

```{r vis-intersects}
sf_y %>% 
  mutate(intersects = st_intersects(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = intersects), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_intersects")
```

```{r vis-disjoint}
sf_y %>% 
  mutate(disjoint = st_disjoint(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = disjoint), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_disjoint")
```


```{r vis-touches}
sf_y %>% 
  mutate(touches = st_touches(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = touches), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_touches")
```


```{r vis-crosses}
sf_y %>% 
  mutate(crosses = st_crosses(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = crosses), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_crosses")
```


```{r vis-within}
sf_y %>% 
  mutate(within = st_within(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = within), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_within")
```

```{r vis-contains}
sf_y %>% 
  mutate(contains = st_contains(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = contains), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_contains")
```


```{r vis-overlaps}
sf_y %>% 
  mutate(overlaps = st_overlaps(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = overlaps), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_overlaps")
```


```{r vis-equals}
sf_y %>% 
  mutate(equals = st_equals(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = equals), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_equals")
```


```{r vis-covers}
sf_y %>% 
  mutate(covers = st_covers(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = covers), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_covers")
```

```{r vis-covered_by}
sf_y %>% 
  mutate(covered_by = st_covered_by(., b0, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = covered_by), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_covered_by")
```

```{r vis-equals_exact}
sf_y %>% 
  mutate(equals_exact = st_equals_exact(., b0, par = 0.01, sparse = FALSE)) %>% 
  
  ggplot() +
  geom_sf(aes(color = equals_exact), fill = NA) +
  geom_sf(data = b0, color = "black", fill = NA) +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "st_equals_exact")
```




