---
title: "Geospatial Data"
author: "Zach"
date: 2020-01-20
output:
  github_document
---

Recipe for plotting geospatial data. Most of the difficulty is in setting up
your environment; using `sf` once installed is relatively straightforward.

## Environment Setup
<!-- -------------------------------------------------- -->

I find that manually installing the correct dependencies for the `sf` package is
nightmarish. Instead, I use the
[rocker/geospatial](https://hub.docker.com/r/rocker/geospatial) Docker image to
automatically handle this installation. Most of the difficulty in environment
setup is then due to setting up docker.

Idea: Run a docker container with all dependencies. Writing geospatial analyses
can then be done through a web browser accessing the docker container.

Here are the steps:

1. Install [docker](https://docs.docker.com/install/)
2. Copy dependencies to target folder:
   a. `Makefile` - automates setup and running
   b. `run-rstudio` - automates starting docker
   c. `user-settings` - settings file for Rstudio
3. Run `$ make setup`. This will pull `rocker/geospatial` and make a directory for data
4. Add `export RSTUDIO_PASS=USER_SELECTED_PASSWORD` to your terminal config. Use this to login to the Rstudio instance in the container.

The steps above *set up* the container. To activate, run `$ make start`; this
should open a web browser. The login info will be

> Username: rstudio
> Password: USER_SELECTED_PASSWORD

You may need to manually open your browser, in which case navigate to
`http://localhost:8787/`. When you're done working, you can run `$ make stop` to
halt the Docker container.

## Code Setup
<!-- -------------------------------------------------- -->

```{r setup}
library(tidyverse)
library(sf)
library(lubridate)
```

## Downloading data
<!-- -------------------------------------------------- -->

With the data directory mounted, you can download data programmatically and extract to the directory. This is useful if you want to document where the data came from, for reproducibility.

As an example, I found Australia fire reports and boundary data.

```{r download}
# Latests Australia fire data
url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"
df_aus_fires <- sf::st_read(url)

# Australia boundary data
url_aus_geo <- "http://data.daff.gov.au/data/warehouse/nsaasr9nnd_022/nsaasr9nnd_02211a04es_geo___.zip"
curl::curl_download(url_aus_geo, destfile = "../data/aus.shp.zip")
unzip("../data/aus.shp.zip", exdir = "../data")
```

Load boundary data.

```{r boundaries}
df_boundaries <- sf::read_sf("../data", "aust_cd66states")
```

## Wrangle
<!-- -------------------------------------------------- -->

Simple feature `sf` data can be used in a Tidyverse pipeline. We can use this to process the descriptions for `control` information.

The fire data come with prose descriptions of fires:

```{r view-control}
df_aus_fires %>% 
  head(2) %>% 
  pull(description)
```

This is human-readable, but not machine-readable. We can extract the `STATUS` information with a regular expression:

```{r proc_control}
df_aus_control <-
  df_aus_fires %>%
  mutate(
    control = str_extract(description, "STATUS: [\\w|\\s]+") %>%
      str_remove(., "STATUS: ") %>%
      str_trim(.) %>%
      ordered(., c("Out of control", "Being controlled", "Under control")),
    pubDate = dmy_hms(pubDate)
  )

df_aus_control %>% glimpse
```

Check the levels of the `control` data:

```{r}
df_aus_control %>% pull(control) %>% unique
```

## Visualize
<!-- ----------------------------------------------------------------- -->

### GIS
<!-- --------------------------- -->

Geographic distribution of fires and their degree of control. 

```{r gis_vis}
fires_crs <- st_crs(df_aus_control)

# Lat-lon data for Australian cities
df_cities_latlon <- tribble(
       ~lat,     ~lon, ~city,
  +115.8605, -31.9505, "Perth",
  +138.6007, -34.9285, "Adelaide",
  +153.0251, -27.4698, "Brisbane",
  +151.2093, -33.8688, "Sydney",
  +149.1300, -35.2809, "Canberra",
  +144.9631, -37.8136, "Melbourne"
)
sf_cities_latlon <-
  st_as_sf(
    df_cities_latlon,
    coords = c("lat", "lon"),
    crs = 4326,
    agr = "constant"
  )

ggplot() +
  geom_sf(data = df_boundaries %>% st_set_crs(fires_crs)) +
  geom_sf(
    data = df_aus_control,
    mapping = aes(fill = control, color = control)
  ) +
  geom_sf(
    data = sf_cities_latlon,
    shape = 21,
    size = 3,
    color = "red",
    fill = "black"
  ) +
  geom_text(
    data = df_cities_latlon,
    aes(lat, lon, label = city),
    hjust = 0,
    nudge_x = +0.5
  ) +
  ## Tail
  scale_fill_discrete(name = "Fire control status") +
  coord_sf(
    xlim = c(+142, +155),
    ylim = c(-38, -27)
  ) +
  guides(color = FALSE) +
  theme(legend.position = "right") +
  theme_minimal() +
  labs(x = "", y = "", caption = str_c("Current as of: ", today()))
```

The fires are concentrated around Canberra and Sydney. Of course, smoke is probably affecting nearby cities as well.

