---
title: "Wrangling Excel Documents"
author: "Zach"
date: 2019-12-10
output:
  github_document
---

**Purpose**: Wrangling Excel files is often *nightmarish*. This notebook details
some tricks I've learned to handle issues found in the wild.

```{r setup}
library(tidyverse)
library(rzdr)
library(readxl)
library(httr)

## Use my tidy-exercises copy of UNDOC data for stability
url_undoc <- "https://github.com/zdelrosario/tidy-exercises/blob/master/2019/2019-12-10-news-plots/GSH2013_Homicide_count_and_rate.xlsx?raw=true"
```

## Example: UNDOC Homicide Dataset
<!-- -------------------------------------------------- -->

The following code will use `httr` to access the Excel file at `url_undoc`, then
pass to `read_excel`.

```{r undoc-load}
GET(url_undoc, write_disk(tf <- tempfile(fileext = ".xlsx")))

df_raw <- read_excel(
  tf,
  sheet = 1,
  skip = 6,
  col_names = c(
    "region",
    "sub_region",
    "territory",
    "source",
    "org",
    "indicator",
    "2000",
    "2001",
    "2002",
    "2003",
    "2004",
    "2005",
    "2006",
    "2007",
    "2008",
    "2009",
    "2010",
    "2011",
    "2012"
  )
)
```

Let's take a look at the head.

```{r undoc-head}
df_raw %>% head
```

We'll handle the implicit values with *lag-filling*.

Let's also inspect the tail.

```{r undoc-tail}
df_raw %>% tail(20)
```

We'll handle the notes in the `region` column with an `NA` replacement.

### Sheet notes
<!-- ------------------------- -->

**Issue**: A minor annoyance in these sheets is the addition of footnotes. These
are important for understanding the data and its caveats, but they make the data
less machine-readable.

**Solution**: I inspected the notes and looked for keywords. I then use an
`if_else` call to replace those instances with missing values. Remember that
`NA_character_` is the correct missing value for character-type vectors.

```{r undoc-rm-notes}
## Detect notes and fill with NA
df_no_notes <-
  df_raw %>%
  mutate(
    region = if_else(str_detect(region, "estimate|data"), NA_character_, region)
  )
```

### Lag-filling
<!-- ------------------------- -->

**Issue**: One of the most obnoxious issues with Excel sheets is that important
labeling entries are often left blank: These values are implicit by location.
While fine for human-readable formats, these implicit values need to be
made explicit for a machine-readable format.

**Solution**: We will repeatedly fill the implicit columns with lagged values.

```{r undoc-wrangle}
## Choose variables to lag-fill
vars_lagged <- c("region", "sub_region", "territory", "source", "org")

## Trim head and notes
df_filled <-
  df_no_notes %>%
  slice(-1) %>%
  slice(-(n()-5:-n()))

## Helper function to count num rows w/ NA in vars_lagged
countna <- function(df) {
  df %>%
    filter(rowAny(across(all_of(vars_lagged), is.na))) %>%
    dim %>%
    .[[1]]
}

## Repeatedly lag-fill until NA's are gone
while (countna(df_filled) > 0) {
  df_filled <-
    df_filled %>%
    mutate(across(
      all_of(vars_lagged),
      ~if_else(is.na(.) & !is.na(lag(.)), lag(.), .)
    ))
}
```

### Tidying
<!-- ------------------------- -->

The final steps are fairly standard wrangling; ensure all the years are integer
values and pivot to tidy the data.

```{r undoc-cast}
## Cast and pivot
df_tmp <-
  df_filled %>%
  mutate(across(`2000`:`2012`, as.numeric)) %>%
  pivot_longer(
    cols = `2000`:`2012`,
    names_to = "year",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = indicator,
    values_from = "value"
  )

## Clean names
df_undoc <-
  df_tmp %>%
  rename_all(str_to_lower)
```

Let's check some summaries.

```{r undoc-summary}
df_undoc %>% summary
```
