---
title: "GGpairs"
author: "Zach"
date: 2019-12-30
output:
  github_document
---

GGpairs helps one create [scatterplot
matrices](https://en.wikipedia.org/wiki/Scatter_plot#Scatter_plot_matrices), and
is provided by the packages `GGally`.

```{r setup}
library(tidyverse)
library(GGally)
```

Basic usage; plot a few variables.

```{r vis-basic}
mpg %>%
  select(year, cyl, drv) %>%
  ggpairs
```

When plotting many variables, the printout can get annoying. Use `print()` to
suppress.

```{r vis-select}
mpg %>%
  select_if(is.numeric) %>%
  ggpairs %>%
  print(progress = F)
```

`ggpairs` can take an `aes()` specification. Make sure you don't select out the
column you're trying to visualize!

```{r vis-aes, fig.width=12, fig.height=10}
mpg %>%
  select(displ, year, cyl, drv, cty, hwy, class) %>%
  ggpairs(., mapping = aes(color = class)) %>%
  print(progress = F)
```

Changing the other figure aesthetics is a bit more involved. Keywords `lower`
and `upper` are used to target the above- and below-diagonal plot components.
For instance, we can add smoothing to the lower section:


```{r vis-lower-smooth, fig.width=12, fig.height=10}
mpg %>%
  select(displ, year, cyl, drv, cty, hwy, class) %>%
  ggpairs(
    .,
    lower = list(continuous = wrap("smooth"))
  ) %>%
  print(progress = F)
```

We can also change the point sizes through the lower keyword:

```{r vis-lower-point-size, fig.width=12, fig.height=10}
mpg %>%
  select(displ, year, cyl, drv, cty, hwy, class) %>%
  ggpairs(
    .,
    lower = list(continuous = wrap("points", size = 2))
  ) %>%
  print(progress = F)
```
