---
title: "Plotly Conversion"
author: "Zach"
date: 2020-12-28
output: html_document
---

[Plotly](https://plotly.com/) is a really slick interactive graphics frontend, but it uses a different interface than ggplot. Thankfully there's a *trivially easy way* to convert a ggplot into a plotly figure!

```{r setup}
library(tidyverse)
library(plotly)
```

Converting a ggplot object to a plotly object is easy with `ggplotly()`. Note that we can add to the hover text in plotly by setting the `text` aesthetic.

```{r small-dataset}
p1 <- 
  mtcars %>% 
  as_tibble(rownames = "model") %>% 
  ggplot(aes(wt, mpg, text = model)) +
  geom_point()
ggplotly(p1)
```

In order to use interactivity in a knitted document, you need to knit to `html` format.

Large datasets can be slow with vanilla `ggplotly()`. An additional call of `toWebGL()` gives a more scalabale graphic.

```{r large-dataset}
p2 <- 
  diamonds %>% 
  ggplot(aes(carat, price, color = cut)) +
  geom_point()
toWebGL(ggplotly(p2))
```

With plotly you can click legend entries to toggle layers of the graphic; e.g. you can click `Ideal` to remove the `Ideal` cut diamonds to help see occluded points.
