---
title: "`ggside` sidebar tools"
author: "Zach"
date: 2019-12-10
output:
  github_document
---

**Purpose**: A reminder on how to use GGside to add sidebar plots

GGside is a useful package for adding summary data to the sides of a plot. Here are some examples, adapted from the [vignette](https://cran.r-project.org/web/packages/ggside/vignettes/ggside_basic_usage.html?utm_source=pocket_mylist).

Available geoms (xside and yside each):
- Bar
- Boxplot
- Density
- Freqpoly
- Histogram
- Line
- Path
- Point
- Text
- Tile
- Violin

# Setup

```{r setup}
library(tidyverse)
library(ggside)
library(palmerpenguins)
```

Process a dataset to visualize.

```{r}
df_diamonds_summary <- 
  diamonds %>%
  mutate(`Cut Clarity` = paste(cut, clarity)) %>%
  group_by(`Cut Clarity`,cut, clarity, color) %>%
  summarise(
    n = n(),
    price_mean = mean(price),
    price_sd = sd(price)
  )
df_diamonds_summary
```

# Recipies

## Tile plots

Visualize with tile sides:

```{r vis-tiles}
df_diamonds_summary %>% 
  ggplot(aes(color, `Cut Clarity`)) +
  geom_tile(aes(color, fill = price_mean)) +
  geom_xsidetile(
    data = . %>% 
      group_by(color) %>% 
      summarize(val = median(price_mean)),
    mapping = aes(xfill = val, y = "median of means")
  ) +
  geom_ysidetile(
    data = . %>% 
      group_by(`Cut Clarity`) %>% 
      summarize(val = median(price_mean)),
    aes(x = "median of means", yfill = val)
  ) +
  theme()
```

*Notes*:

- `ggside` doesn't manually carry out grouping for summaries under-the-hood; you'll likely have to post-process the `data` within your plot.
- `ggside` functions need manual overrides for the unused axis
  - for instance, if I didn't set `x` in the `ysidetile` call above, I'd get a repeated `color` axis
  
## Density plots

```{r vis-densities}
penguins %>% 
  
  ggplot(aes(bill_length_mm, bill_depth_mm, color = species)) +
  geom_point() +
  geom_xsidedensity() +
  geom_ysidedensity() +
  theme(legend.position = "bottom")
```

*Notes*

- The relevant vignette manually sets the density aesthetic, but `ggside` handles this automatically; you can leave out the `aes` call
