---
title: "Marginals with ggExtra"
author: "Zach"
date: 2019-12-25
output:
  github_document
---

The package `ggExtra` has tools to add marginal distributions to a `ggplot`.
However, the syntax is non-standard, and does not fit in the usual Tidyverse
pipeline.

```{r setup}
library(tidyverse)
library(ggExtra)
```

To use `ggExtra::marginals()`, we have to pass the plot itself to the function.
I find that piping doesn't work; I have to save to an intermediate variable and
pass explicitly:

```{r baseline}
p <-
  mpg %>%
  pivot_longer(names_to = "type", values_to = "mpg", cty:hwy) %>%

  ggplot(aes(displ, mpg)) +
  geom_point()

ggMarginal(p)
```

You can change the type of marginal with the `type` keyword. Valid choices
are `density, histogram, boxplot, violin, densigram`.

```{r densigram}
p <-
  mpg %>%
  pivot_longer(names_to = "type", values_to = "mpg", cty:hwy) %>%

  ggplot(aes(displ, mpg)) +
  geom_point()

ggMarginal(p, type = "densigram")
```

In order to show groups, you have to enable a flag based on the chosen
aesthetic. This is really obnoxious and easy to forget.

```{r color-groups}
p <-
  mpg %>%
  pivot_longer(names_to = "type", values_to = "mpg", cty:hwy) %>%

  ggplot(aes(displ, mpg, color = type)) +
  geom_point()

ggMarginal(p, groupColour = TRUE)
```
