---
title: "ggrepel"
author: "Zach"
date: 2019-12-31
output:
  github_document
---

**Purpose**: Reminder on key `ggrepel` features.

```{r setup}
library(tidyverse)
library(ggrepel)
```

`ggrepel` provides extremely useful tools for plotting text that automatically
dodges datapoints.

```{r }
tribble(
  ~dtl, ~gam, ~gam_lo, ~gam_up, ~label,
     0,    0,       0,       0, "Perfect",
    -1,   -1,      -1,      -1, "Anti-conservative",
    +1,   +1,      +1,      +1, "Conservative",
    +1,   -1,      -1,      -1, "False detection",
    -1,   +1,      +1,      +1, "Missed detection",
  +0.5, +0.5,      -1,      +2, "Inconclusive"
) %>%

  ggplot(aes(dtl, gam)) +
  geom_point() +
  geom_text_repel(aes(label = label))
```

However, the settings can be rather overwhelming. Key arguments are summarized
in the [vignette](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html):

| Option | Default| Description |
|---|---|
| force | 1 | force of repulsion between overlapping text labels |
| direction | "both"| move text labels “both” (default), “x”, or “y” directions |
| max.iter | 2000 | maximum number of iterations to try to resolve overlaps |
| nudge_x | 0 | adjust the starting x position of the text label |
| nudge_y | 0 | adjust the starting y position of the text label |
| box.padding | 0.25 lines | padding around the text label |
| point.padding | 0 lines | padding around the labeled data point |
| segment.color | "black" | line segment color |
| segment.size | 0.5 mm | line segment thickness |
| segment.alpha | 1.0 | line segment transparency |
| arrow | NULL | render line segment as an arrow with grid::arrow() |

One of the *most* key arguments is `point.padding`, which helps prevent text
from overlapping the points:

```{r }
tribble(
  ~dtl, ~gam, ~gam_lo, ~gam_up, ~label,
     0,    0,       0,       0, "Perfect",
    -1,   -1,      -1,      -1, "Anti-conservative",
    +1,   +1,      +1,      +1, "Conservative",
    +1,   -1,      -1,      -1, "False detection",
    -1,   +1,      +1,      +1, "Missed detection",
  +0.5, +0.5,      -1,      +2, "Inconclusive"
) %>%

  ggplot(aes(dtl, gam)) +
  geom_point() +
  geom_text_repel(
    aes(label = label),
    point.padding = 1
  )
```
