---
title: "Scraping an HTML Table"
author: "Zach"
date: 2020-01-19
output:
  github_document
---

Simple recipe for scraping HTML tables.

```{r }
library(tidyverse)
library(rvest)
library(lubridate)

str_to_var <- function(str) {
  str %>%
    str_trim(.) %>%
    str_to_lower(.) %>%
    str_replace_all("\\s+", "_")
}
```

## Page Reading and Selecting
<!-- -------------------------------------------------- -->

First, we need the URL of the target page

```{r url}
url_page <- "https://en.wikipedia.org/wiki/Oldest_people"
```

Next, we need a *css selector* for our chosen table. To do so, we open the
target page in our browser, open the developer tools, find the html element we
want to select, and copy the CSS selector.

In Firefox Option + Command + C brings open the relevant console.

Doing the previous on the Oldest people page yields the following two selectors.

```{r selector}
css_women <- "table.wikitable:nth-child(15)"
css_men <- "table.wikitable:nth-child(18)"
```

## Wrangle
<!-- -------------------------------------------------- -->

We use rvest to read html pages and use the selectors to
grab certain tables.

```{r read}
## Read the page based on url
page <- read_html(url_page)

## Use html_nodes() and html_table()
df_women <-
  page %>%
  html_nodes(css_women) %>%
  html_table() %>%
  .[[1]] %>%
  as_tibble() %>%

  mutate(sex = "F") %>%
  rename_all(~str_remove_all(., "\\[.*\\]")) %>%
  rename_all(str_to_var)

df_men <-
  page %>%
  html_nodes(css_men) %>%
  html_table() %>%
  .[[1]] %>%
  as_tibble() %>%

  mutate(sex = "M") %>%
  rename_all(~str_remove_all(., "\\[.*\\]")) %>%
  rename_all(str_to_var)

## Bind two tables
df_data <-
  bind_rows(
    df_women,
    df_men
  ) %>%
  mutate(
    name = str_remove_all(name, "\\[.*\\]") %>%
        fct_reorder(., age),
    age = str_extract(age, "\\d+") %>% as.integer,
    birth_date = dmy(birth_date),
    death_date = dmy(death_date) %>%
      if_else(is.na(.), today(), .)
  )
```

## Visualize
<!-- -------------------------------------------------- -->

```{r vis}
df_data %>%

  ggplot() +
  geom_segment(
    aes(x = birth_date, xend = death_date, y = name, yend = name, color = sex)
  ) +

  scale_color_discrete(name = "") +
  theme_minimal() +
  labs(
    x = "",
    y = "Person"
  )

```
