---
title: "TidyEval: Brace Syntax"
author: "Zach"
date: 2020-02-21
output:
  github_document
---

The syntax for tidy evaluation changed (Circa Feb. 2020) since I first learned
it in the DCL. The syntax now uses braces and dots rather than explicit
quotation and dereferencing. The page [Programming with
dplyr](https://dplyr.tidyverse.org/dev/articles/programming.html) has full
up-to-date details on this syntax, but I'm recording the quotation stuff here for
my own reference.

```{r library}
library(tidyverse)

```

When writing a function that takes **tibble variable names** (as strings) and
uses dplyr operations, you can use the `.data` pronoun with square braces `[[
]]`.

```{r square-brace-function}
for (var in names(mtcars)[1:2]) {
  mtcars %>%
    count(.data[[var]]) %>% print()
}
```

When writing a function that takes **tibble variables** and uses dplyr operations,
you need to use curly braces `{{ }}` to signal this *indirection*.

```{r curly-brace-function}
var_summary <- function(data, var) {
  data %>%
    summarize(
      n = n(),
      min = min({{ var }}),
      max = max({{ var }})
    )
}

mtcars %>%
  group_by(cyl) %>%
  var_summary(mpg)
```

To capture **multiple tibble variables**, use the `...` pattern. When using this
feature, the [Tidyverse style
guide](https://design.tidyverse.org/dots-prefix.html) recommends dot-prefixing
all other argument names. This is to help prevent unintentional capture in the
`...` set.

```{r dot-function}
my_summarize <- function(.data, ...) {
  .data %>%
    group_by(...) %>%
    summarize(
      mass = mean(mass, na.rm = TRUE),
      height = mean(height, na.rm = TRUE)
    )
}

starwars %>% my_summarize(species, gender)
```
