---
title: "Test: Support Points"
author: "Zach"
date: 2020-10-08
output:
  github_document
---

[Support points](https://arxiv.org/abs/1609.01811) allow one to sample a low-discrepancy means to "compact" a given distribution---either analytic or empirical.

- The `support` [package](https://cran.r-project.org/web/packages/support/index.html) was removed from CRAN; archived versions can be installed via source. I use version 0.1.4 in this notebook.

The purpose of this notebook is to test support points as a means for design-of-experiments (DOE) given a target distribution.

```{r setup}
library(tidyverse)
library(support) # Not on CRAN; manual install required
library(mvtnorm)
```

Support points can be generated to "compact" a given dataset; let's test that functionality.

```{r test}
## Generate target points
X_norm <- rmvnorm(
  n = 1000,
  sigma = matrix(c(1, 0.5, 0.5, 1), nrow = 2)
)

## Generate support points
res <- sp(
  n = 50, # Samples
  p = 2,  # Target dimensionality
  dist.samp = X_norm
)

## Extract the support points
X_sp <- res$sp
```

Visualize the points

```{r vis}
df <-
  as_tibble(X_norm) %>%
  mutate(source = "original") %>%
  bind_rows(
    as_tibble(X_sp) %>%
    mutate(source = "sp")
  )

df %>%
  ggplot(aes(V1, V2, color = source)) +
  geom_point()
```
